\documentclass[11pt,oneside]{amsart}
\usepackage{geometry}
\usepackage{amssymb,parskip}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}

\title{MATH1103 Fall 2022\\
Discussion Sheet 10}

\begin{document}
    \maketitle
 
    \begin{problem} Hard problems on Problem Set 08.
    \newline
    Use Zax Theorem to show that the series
    $$
    z_n=\frac{1}{2}-\frac{1}{3}+\cdots+\frac{(-1)^n}{n}
    $$
    converges.
    \begin{enumerate}
        \item What is the definition of a Zax pair? What is the statement of Zax Theorem?
        \vfill
        \item What is the Zax pair in this question in particular? Then write down the proof.
        \vfill
        \end{enumerate}
    \end{problem}
\newpage

    \begin{problem}
    \textit{(Comparison Test)}: Determine whether the following series diverges or converges.
    \begin{enumerate}
    \item 
    $$
    \sum_{n=1}^\infty \dfrac{n}{n^3+3}
    $$
    \vfill
    \item
    $$
    \sum_{n=1}^\infty \dfrac{3}{n\cdot {\sin}^2 n}
    $$
    Note: You may use the fact that $\sin n \not =0,$ for any positive integer $n$.
    \vfill
    \end{enumerate}
    \end{problem}
    \newpage
    \begin{problem}\textit{Integral Test}.
    The Integral test for convergence of a series states as follows (Let $a_n=f(n)$ for all positive integer $n$, where $f(x)$ is a function):
    $$
    \text{If }\int_1^\infty f(x) \text{ }dx \text{ converges, then }\sum_{n=1}^\infty a_n \text{ converges}.
    $$
    $$
    \text{If }\int_1^\infty f(x) \text{ }dx \text{ diverges, then }\sum_{n=1}^\infty a_n \text{ diverges}.
    $$
    \begin{enumerate}
        \item Recall that the harmonic series
        $$
        \sum_{n=1}^\infty \frac{1}{n}
        $$
        diverges. What kind of technique did we use last time? Indeed, the divergence of harmonic series can also be proved by the integral test!
        \vfill
        \item Consider the series
        $$
        \sum_{n=1}^\infty \frac{1}{n^2}
        $$
        By integral test, determine whether this diverges or converges.
        \vfill
        \item Based on the last two parts, can you generalize the idea of the proof to the general statement as shown at the beginning of the questions?
    \end{enumerate}
    \end{problem}

\end{document}