\documentclass[11pt,oneside]{amsart}
\usepackage{geometry}
\usepackage{amssymb,parskip}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}

\title{MATH1103 Fall 2022\\
Discussion Sheet 4}

\begin{document}
    \maketitle
    Every discussion you will be assigned three problems and you are encouraged to work in groups.

    \begin{problem}
         If a bus arrives at a random time between 10 am and 10:30 am, and you arrive at a random time between 10:15 am and 10:30 am, what's the probability of your missing the bus?
    \end{problem}

    \begin{problem}
       This problem aims to help with practicing the methods of u-substitution and integration by parts. Calculate the following integrals:
       \begin{enumerate}
           \item $$\int_0^1 \frac{x^3}{(1+x^4)^{\frac{1}{3}}} dx$$
           \item $$ \int_0^1 (x+1)e^{x^2+2x+3} dx$$
           \item $$\int xe^x dx$$
       \end{enumerate}
       
       
       
    \end{problem}

    \begin{problem}
       This problem aims to give you some guidance towards the last problem of your homework this time. In the homework, you are asked to calculate the volume of a 4D sphere... Okay, it sounds horrible -- but let's start with lower dimensional cases so that hopefully you can tackle the higher dimensional ones!
       \begin{enumerate}
           \item What is a 1D sphere? What is a 2D sphere? What is a 3D sphere?
           \item Can you slice a 2D sphere into different pieces of 1D spheres?(It could be infinitely many pieces)
           \item After the slicing, how could you calculate the "volume" of a 2D sphere?
           \item Can you do the same thing to a 3D sphere?
           
       \end{enumerate}
        
    \end{problem}


\end{document}