\documentclass[11pt,oneside]{amsart}
\usepackage{geometry}
\usepackage{amssymb,parskip}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}

\title{MATH1103 Fall 2022\\
Discussion Sheet 5}

\begin{document}
    \maketitle
    Every discussion you will be assigned three problems and you are encouraged to work in groups.

    \begin{problem} This is a very open-ended problem! Feel free to share your ideas!
        \begin{enumerate}
         \item What are the formulas for the circumference and the area of a circle? Any relationship between these two formulas observed? Is that a coincidence?
         \item What are the formulas for the surface area and the volume of a sphere? Any relationship? Is that a coincidence as well?
         
         \end{enumerate}
    \end{problem}

    \begin{problem}
       Let's say we roll \textbf{two} ordinary dices simultaneously. Then we can denote the outcome as a pair, i.e., $(x,y)$ where $x$ is the outcome for one dice while $y$ for the other one. Answer the following questions accordingly:
       \begin{enumerate}
           \item What is the probability of getting a doublet? A doublet looks like $(a,a)$.
           \item If we consider the sum of the pair, i.e., $x+y$, what is the probability of $x+y=8$? What about $x+y \leq 6$?
           \item If a normal dice is thrown $5$ times, what is the probability of getting at least three consecutive $6$'s?
       \end{enumerate}
       
       
       
    \end{problem}

    \begin{problem}
       Jack bought 5 pairs of socks last weekend but he was too lazy to separate each of the pairs. Starting from next Monday, he chooses two random socks every morning without washing them until the end of the week. Answer the following questions:
       \begin{enumerate}
           \item What is the probability that he finds a matching pair on Monday morning?
           \item What is the probability that he finds a matching pair on Tuesday morning? What about Wednesday, Thursday, Friday?
            
       \end{enumerate}
        
    \end{problem}


\end{document}