\documentclass[11pt,oneside]{amsart}
\usepackage{geometry}
\usepackage{amssymb,parskip}
\usepackage[shortlabels]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}

\newcommand{\bC}{\mathbb{C}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bZ}{\mathbb{Z}}

\title{MATH1103 Fall 2022\\
Discussion Sheet 8}

\begin{document}
    \maketitle

    \begin{problem}
    Recall that a sequence $\{a_n\}$ converges to $L$ if for any $\epsilon >0$, there exists a positive integer $N$ such that 
    $$
    |a_n -L| <\epsilon, \text{ for all }n \geq N
    $$
    Then show that $a_n=\frac{1}{n^2} \to 0$ by the $\epsilon-$definition of the convergence of a sequence.
    \vfill
    \end{problem}
    
    \begin{problem}
    During the class yesterday, you learned that if $\{x_n\} \to x$ and $\{y_n\} \to y$, then $\{x_n+y_n\} \to x+y$ and $\{x_ny_n\} \to xy$. Similarly, we can ask questions as follows: If $\{x_n\}$ and $\{y_n\}$ both diverge, then must $\{x_n+y_n\}$ and $\{x_ny_n\}$ diverge as well? If yes, prove it! If not, give counterexamples! 
    \newline
    \textit{Hint}: Can you give some examples of divergent sequences? 
    \vfill
    \end{problem}
 
 \newpage   
    \begin{problem}\textit{(Geometric series)}
    Let's work with our old friend again! Let $a$ be any real number, and $r$ be a positive real number.
    \begin{enumerate}
        \item Consider the finite series as follows. 
        $$
        s_n=a+ar+ar^2+ar^3+\cdots+ar^n
        $$
        How can you calculate such a sum?
        \vfill
        \item Then let's consider the infinite series as follows.
        $$
        s=a+ar+ar^2+ar^3+\cdots
        $$
        For which $r's$ is this series convergent? For the convergent case, how can you calculate $s$? Can you use some similar argument to the first part?
        \vfill
        \item Prove that $1=.999\cdots$ following the idea of geometric series.
    \end{enumerate}
    \end{problem}

    


\end{document}