\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
\usepackage{amsmath}								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{}
\author{MATH1103   \ \ \ \ \ \ Midterm 3 \ \ \ \  November 19, 2021 \ \ \ \ \ \ Version A \\ Closed book, closed notes; show all your work.}
\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}
\begin{enumerate}

\item (12 pts) For each of the following, tell if it's True or False; \emph{no reasons need to be given.}
\begin{enumerate}
\item(T/F): If $f$ is any continuous function, then $$\int_{-\infty}^{\infty} f(x) \, dx=\lim_{t \to \infty} \int_{-t}^t f(x) \, dx.$$

\

\item
T/F: if $\lim_{n \to \infty}a_n=L$, then $\lim_{n \to \infty}a_{2n+1}
=L$ 

\

\item
T/F: if $\{a_n\}$ is decreasing and $a_n>0$ for all $n$, then $\{a_n\}$ must converge.
\end{enumerate} 

\

\item (18 pts) Evaluate the following integrals if they converge or show they diverge, showing all your work.
$$(a) \ \int_{1}^{\infty} \dfrac{x}{e^{2x}} \, dx \ \ \ \ \ \ \ \ \ \ \ (b) \int_0^6 \dfrac{x}{x^2-4} \, dx $$

\



\item (18 pts) Find the value of the following converging series (you don't have to show they converge):

$$(a) \ \sum_{k=2}^{\infty} \left( \dfrac{1}{k^{3}}- \dfrac{1}{(k+2)^{3}} \right) \ \ \ \ \ \ \ (b) \sum_{n=1}^{\infty} \dfrac{3^{n-4}}{5(4^n)} $$

\

\item (10 pts) Use the integral test to determine if the following series converges or diverges. \emph{You do not need to show that it satisfies the hypotheses for the test.}

$$ \ \sum_{n=1}^{\infty} \dfrac{6n^2}{4n^3+1} $$

\item (18 pts) Use the comparison or limit comparison test to determine if the following series converge or diverge:

$$(a) \ \sum_{n=3}^{\infty} \dfrac{1}{n\sqrt{2n^2-4}} \ \ \ \ \ \ \ \ (b) \ \sum_{n=1}^{\infty} \dfrac{2+\cos(n^2)}{n^3} $$

\

\item (10 pts) The alternating series $\sum_{n=0}^{\infty} (-1)^n/n!$ is easily shown to converge.  (You don't need to prove this.) We will later show that it converges to $1/e$. Estimate $1/e$ to within .0001, using the error bound for an alternating series. (That is, find the minimum $n$ so that $R_n<.0001.$ Then find $s_n$ for that choice of $n$. You need a numerical answer, but don't need to compute factorials or simplify your answer.)

\

\item (14 pts) Two sequence problems.

\begin{enumerate} 



\item Suppose a sequence $\{ a_n \}$ satisfies $c_1=2$ and, for $n \ge 1$, $c_{n+1}=5+3/c_n$. If the sequence converges, what does it converge to?

\

\item Consider the sequence $$\{a_n\}=\left\{\dfrac{n}{n+5} \right\}_{n=6}^{\infty}.$$ By computing $a_{n+1}-a_n$, determine if the sequence is increasing, decreasing, or neither. \emph{Using the monotone sequence theorem} (and justifying its use) determine if the sequence converges. You \emph{must} use the monotone sequence theorem for full credit.



\end{enumerate}



\end{enumerate}
\end{document}  